Supplement to `\ourlongtitle'

### 2021-10-13 Final_version/glo-2020-theor-results.tex
Added final version

In \cite{olsson:westerborn:2014b}, the authors established strong consistency and asymptotic normality of the PaRIS in the framework of fully dominated general state-space HMMs and the bootstrap particle filter, \ie, in the simple case where $\adjfuncforward{n} \equiv 1$ and $\kissforward{n}{n} \equiv \hd{n}$ for all $n$. In the following we will extend these results to the considerably more general setting comprising models \eqref{eq:def:post} and the pseudo-marginal PaRIS in Algorithm~\ref{alg:pm:PaRIS}. More precisely, we will show that each weighted sample $(\epart{n}{i}, \tstat[i]{n}, \ewght{n}{i})_{i = 1}^N$, $n \in \nset$, produced by Algorithm~\ref{alg:pm:PaRIS} satisfies exponential concentration (Theorem~\ref{cor:hoeffding:tau:marginal}) and asymptotic normality (Theorem~\ref{cor:clt:pseudo:marginal:paris}) with respect to the expected additive functional $h_n$ under the \emph{skew} path model 
$$
\postmod{0:n}(\rmd x_{0:n}) \propto \chi(\rmd x_0) \prod_{m = 0}^{n - 1} \ukmod{m}(x_m, \rmd x_{m + 1}), \quad n \in \nset. 
$$
Even though these results are established along the lines of the corresponding proofs in \cite{olsson:westerborn:2014b}, it is a matter of non-trivial adaptations. As explained in Section~\ref{sec:introduction}, a random-weight particle filter (iterated pseudo-marginal forward sampling) can be viewed as a standard particle filter evolving on an extended state space comprising also the states of the auxiliary variables, and hence its convergence follows from standard SMC convergence results \cite{fearnhead2008particle}. On the contrary, since Algorithm~\ref{alg:pm:PaRIS} involves \emph{two} levels of pseudo marginalisation (with respect to both the forward-sampling and the backward-sampling operations), it cannot be described equivalently as a special instance of the original PaRIS (even in its general form given by Algorithm~\ref{alg:ideal:PaRIS}) operating on an extended space. Thus, there is no free lunch when it concerns the theoretical analysis of this scheme. Furthermore, in the case of fully dominated HMMs and when forward sampling is guided by the bootstrap filter, which was the setting in \cite{olsson:westerborn:2014b}, the conditional distribution of the particles given their ancestors (\ie, the marginal of $\rho_n$ in \eqref{eq:cond:instrumental:mixture} with respect to $x$) coincides, at any time point $n$, with the denominator of the backward kernel. Mathematically, this enables a cancellation that facilitates the analysis significantly. However, this simplification is not possible once the particle dynamics is guided by general proposal kernels, which is necessarily the case in Algorithm~\ref{alg:pm:PaRIS}. This complicates the analysis; see Remark~\ref{rem:non-trivial:extension} in the supplement for further details.

@@

We now focus on the asymptotic properties of Algorithm~\ref{alg:pm:PaRIS} and furnish, in Theorem~\ref{cor:clt:pseudo:marginal:paris} below, this scheme with a CLT. The stochastic stability of the algorithm is expressed by establishing, in Theorem~\ref{thm:variance:bound}, an $\mathcal{O}(n)$ bound on the asymptotic variances. %In order to be able to state these results accurately, we need some additional notation. 
First, let for $(x_n, x_{n + 1}) \in \Xset_n \times \Xset_{n + 1}$, 
\begin{multline} \label{eq:def:ukestvar}
\ukestvar{n}(x_n, x_{n + 1}) \eqdef 
\frac{1}{\wgtfuncmod{n}(x_n, x_{n + 1})} \int \{ \wgtfunc{n}(x_n, x_{n + 1}, z) - \wgtfuncmod{n}(x_n, x_{n + 1}) \}^2 \, \ukdist{n}(x_n, x_{n + 1}, \rmd z)
\end{multline}
denote the conditional relative variance of the (random) weight $\ewght{n + 1}{i}$ given $\epart{n}{\ind{n}{i}} = x_n$ and $\epart{n + 1}{i}= x_{n + 1}$. In addition, for each $n \in \nset$ and $m \in \intvect{0}{n}$, define the kernel  
\begin{equation} \label{eq:def:retrokmodmod}
    \retrokmodmod_{m, n}(x_m', \rmd x_{0:n}) \eqdef \delta_{x_m'}(\rmd x_m) \,  
    \tstatmod{m}(x_m, \rmd x_{0:m - 1})
    \prod_{\ell = m}^{n - 1} \ukmod{\ell}(x_\ell, \rmd x_{\ell + 1})
\end{equation}
on $\Xset_m \times \Xfd^n$ as well as the centered version 
\begin{equation} \label{eq:def:retrokmodmodnorm}
\retrokmodmodnorm_{m, n} h (x_m) \eqdef  \retrokmodmod_{m, n}(h - \postmod{0:n} h)(x_m) 
\end{equation}
on the same space. The following is the main result of this section. 
\begin{theorem} \label{cor:clt:pseudo:marginal:paris}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $\precpar \in \precparsp$, $\K \in \nsetpos$, and $h_n \in \bmaf{\Xfd^n}$, 
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \postmod{0:n} h_n  \right) 
  \dlim \sigma_n(h_n) Z, 
$$
where $Z$ is standard normally distributed and 
\begin{equation} \label{eq:as:var:decomp}
\sigma^2_n (h_n) \eqdef \frac{\chi(\initwgtfunc \retrokmodmodnorm_{0, n} h_n)^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} + \sigma^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) + \sigma^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) 
\end{equation}
and
\begin{multline} \label{eq:non-recursive:as:var}
\sigma^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h_n) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m}(\varphi_m [\retrokmodmodnorm_{m + 1, n} h_n ]^2)}{(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}\\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\postmod{m} \adjfuncforward{m} \postmod{\ell} \ukmod{\ell} \{\bkmod{\ell}(\tstatmod{\ell} h_{\ell} + \addf{\ell} - \tstatmod{\ell + 1} h_{\ell +1})^2 \ukmod{\ell + 1, m}( \bkmod{m} \varphi_m [\ukmod{m + 1, n - 1} \1_{\Xset^n}]^2
)\}}{\K^{m - \ell + 1} (\postmod{\ell} \ukmod{\ell, m - 1} \1_{\Xset^m})(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}
\end{multline}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_\ell : \Xset_\ell \times \Xset_{\ell + 1} \to \rsetnn$. 
\end{theorem}

@@

Our analysis provides, as a by-product, also a CLT for the random-weight particle filter obtained by iterating the forward-sampling operation in Algorithm~\ref{alg:pm:SMC}; indeed, this result, which is similar to \cite[Theorem~3]{fearnhead2008particle}, follows immediately by letting $f_n \equiv 0$ in Theorem~\ref{prop:clt:pseudo:marginal:paris} in the supplement, and we state it below for completeness.

\begin{proposition} \label{prop:clt:pseudo:marginal:filter}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal} and let $(\epart{n}{i}, \ewght{n}{i})_{i = 1}^\N$, $n \in \nset$, be generated by Algorithm~\ref{alg:pm:SMC}. Then for all $n \in \nset$, $\precpar \in \precparsp$, and $h \in \bmf{\Xfd_n}$, 
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} h(\epart{n}{i}) - \postmod{n} h \right) 
  \dlim \tilde{\sigma}_n(h) Z, 
$$
where $Z$ is standard normally distributed and
\begin{equation} \label{eq:as:var:pm:filter}
\tilde{\sigma}^2_n(h) \eqdef \frac{\chi\{ \initwgtfunc \ukmod{0} \cdots \ukmod{n - 1}(h - \postmod{n} h) \}^2}{(\chi \ukmod{0} \cdots \ukmod{n - 1} \1_{\Xset_n})^2} + \tilde{\sigma}^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h) + \tilde{\sigma}^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h) 
\end{equation}
and
\begin{equation} \label{eq:non-recursive:filter:as:var}
\tilde{\sigma}^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m}(\varphi_m [\ukmod{m + 1} \cdots \ukmod{n - 1} (h - \postmod{n} h)
]^2)}{(\postmod{m} \ukmod{m} \cdots \ukmod{n - 1} \1_{\Xset_n})^2} 
\end{equation}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_\ell : \Xset_\ell \times \Xset_{\ell + 1} \to \rsetnn$. 
\end{proposition}

@@

Importantly, Theorem~\ref{cor:clt:pseudo:marginal:paris} also provides, as another by-product, also a CLT for the ideal PaRIS in Algorithm~\ref{alg:ideal:PaRIS}. Indeed, in the case where every $\ud{n}$ is tractable, we may set 
\begin{equation} \label{eq:ideal:case}
\ukest{n}{z}(x_n, x_{n + 1}) = \ud{n}(x_n, x_{n + 1})
\end{equation} 
for all $z$ and define $\ukdist{n}$ arbitrarily. This implies that the relative variance \eqref{eq:def:ukestvar} is identically zero, which eliminates the last term of \eqref{eq:as:var:decomp}. Thus, in this case the asymptotic variance is given by the first two terms of \eqref{eq:as:var:decomp}, but now induced by the original dynamics $(\uk{n})_{n \in \nset}$ (as \eqref{eq:ideal:case} implies that also $\udmod{n} = \ud{n}$). This result is formulated in the following corollary, where we have defined, for each $n \in \nset$ and $m \in \intvect{0}{n}$, the kernel  
\begin{equation} \label{eq:def:retrok}
    \retrok_{m, n}(x_m', \rmd x_{0:n}) \eqdef \delta_{x_m'}(\rmd x_m) \,  
    \tstat{n}(x_m, \rmd x_{0:m - 1})
    \prod_{\ell = m}^{n - 1} \uk{\ell}(x_\ell, \rmd x_{\ell + 1})
\end{equation}
on $\Xset_m \times \Xfd^n$ as well as the centered version 
$$
\retroknorm_{m, n} h (x_m) \eqdef  \retrok_{m, n}(h - \post{0:n} h)(x_m) 
$$
on the same space.

\begin{corollary} \label{cor:clt:ideal:paris}
For each $n \in \nset$, assume that the functions $\adjfuncforward{n}$ and  
$$
\wgtfuncideal{n} : \Xset_n \times \Xset_{n + 1} \ni (x_n, x_{n + 1}) \mapsto \frac{\ud{n}(x_n, x_{n + 1})}{\adjfuncforward{n}(x) \kissforward{n}{n}(x_n, x_{n + 1})} 
$$
are bounded and let $(\epart{n}{i}, \tstat[i]{n}, \ewght{n}{i})_{i = 1}^\N$, $n \in \nset$, be generated by Algorithm~\ref{alg:ideal:PaRIS}. Then for all $n \in \nset$, $\precpar \in \precparsp$, $\K \in \nsetpos$, and $h_n \in \bmaf{\Xfd^n}$, 
$$
\sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \post{0:n} h_n  \right) \dlim \bar{\sigma}_n (h_n) Z, 
$$
where $Z$ is standard normally distributed and
\begin{multline} \label{eq:as:var:ideal:PaRIS}
\bar{\sigma}^2_n (h_n) 
\eqdef \frac{\chi(\initwgtfunc \retroknorm_{0, n} h_n)^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} + \sum_{m = 0}^{n - 1} \frac{\post{m} \adjfuncforward{m} \post{m} \uk{m}(\wgtfuncideal{m} [\retroknorm_{m + 1, n} h_n ]^2)}{(\post{m} \uk{m, n - 1} \1_{\Xset_n})^2} \\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\post{m} \adjfuncforward{m} \post{\ell} \uk{\ell} \{\bkw{\ell}(\tstat{\ell} h_{\ell} + \addf{\ell} - \tstat{\ell + 1} h_{\ell +1})^2 \uk{\ell + 1, m}( \bkw{m} \wgtfuncideal{m} [\uk{m + 1, n - 1} \1_{\Xset^n}]^2
)\}}{\K^{m - \ell + 1} (\post{\ell} \uk{\ell, m - 1} \1_{\Xset_m})(\post{m} \uk{m, n - 1} \1_{\Xset_n})^2}. 
\end{multline}
\end{corollary}

@@

%Having introduced these quantities, we are ready to present the main result of this section.

@@

Under \hypref{assum:strong:mixing} and \hypref{assum:bias:bound} we may establish the next theorem, whose proof is postponed to Section~\ref{sec:proofs} in the supplement.

@@

\begin{remark}
Consider now a parameterised version of the model, where the transition densities $(\ud{n ; \theta})_{k \in \nset}$ are indexed by some parameter $\theta$ belonging to some parameter space $\Theta$ being a subset of $\rset^d$. Assume further that all $(\ud{n ; \theta})_{n \in \nset}$ are differentiable with respect to $\theta$ and such that for all $n \in \nset$, $x_n \in \Xset$, and $h \in \bmf{\Xfd_{n + 1}}$,  
$$
\nabla_\theta \int \ud{n ; \theta}(x_n, x_{n + 1}) h(x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1}) = \int \nabla_\theta \ud{n ; \theta}(x_n, x_{n + 1}) h(x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1})
$$
and $\sup_{\theta \in \Theta} \int |\nabla_\theta \ud{n ; \theta}(x_n, x_{n + 1})| \, \mu_{n + 1}(\rmd x_{n + 1}) \leq c < \infty$ for some positive constant $c$, implying \hypref{assum:bias:bound} in the sense that for all $n \in \nset$, $h \in \bmf{\Xfd_{n + 1}}$, and $x_n \in \Xset$, 
$$
| \uk{n; \theta}h(x_n) - \uk{n; \theta'}h(x_n) | \leq c \| \theta - \theta' \| \| h \|_\infty
$$
(\emph{i.e.}, $\precpar = \| \theta - \theta' \|$ in this case). Finally, assume also that family satisfies \hypref{assum:strong:mixing} uniformly over the parameter space in the sense that for all $n \in \nset$, $(x_n, x_{n + 1}) \in \Xset^2$, and $\theta \in \Theta$, 
$
\udlow \leq \ud{n; \theta}(x_n, x_{n + 1}) \leq \udup. 
$  
Then Theorem~\ref{thm:bias:bound} provides a positive constant $d$ such that for all $n \in \nset$,  
$$
\big| \post{0:n ; \theta} h_n -  \post{0:n ; \theta'} h_n \big|  
\leq d n \| \theta - \theta' \| \sup_{k \in \nset} \| \addf{k} \|_\infty. 
$$
This extends previous results on the uniform continuity of the filter distribution (see, \eg, \cite{papavasiliou:2006,legland:oudjane:2004}) to smoothing of additive state functionals.  
\end{remark}

>>>>

In \cite{olsson:westerborn:2014b}, the authors established strong consistency and asymptotic normality of the PaRIS in the framework of fully dominated general state-space HMMs and the bootstrap particle filter, \ie, in the simple case where $\adjfuncforward{n} \equiv 1$ and $\kissforward{n}{n} \equiv \hd{n}$ for all $n$. In the following we will extend these results to the considerably more general setting comprising models \eqref{eq:def:post} and the pseudo-marginal PaRIS in Algorithm~\ref{alg:pm:PaRIS}. More precisely, we will show that each weighted sample $(\epart{n}{i}, \tstat[i]{n}, \ewght{n}{i})_{i = 1}^N$, $n \in \nset$, produced by Algorithm~\ref{alg:pm:PaRIS} satisfies exponential concentration (Theorem~\ref{cor:hoeffding:tau:marginal}) and asymptotic normality (Theorem~\ref{cor:clt:pseudo:marginal:paris}) with respect to the expected additive functional $h_n$ under the \emph{skew} path model 
$$
\postmod{0:n}(\rmd x_{0:n}) \propto \chi(\rmd x_0) \prod_{m = 0}^{n - 1} \ukmod{m}(x_m, \rmd x_{m + 1}), \quad n \in \nset. 
$$
Even though these results are established along the lines of the corresponding proofs in \cite{olsson:westerborn:2014b}, it is a matter of non-trivial adaptations. As explained in Section~\ref{sec:introduction}, a random-weight particle filter (iterated pseudo-marginal forward sampling) can be viewed as a standard particle filter evolving on an extended state space comprising also the states of the auxiliary variables, and hence its convergence follows from standard SMC convergence results \cite{fearnhead2008particle}. On the contrary, since Algorithm~\ref{alg:pm:PaRIS} involves \emph{two} levels of pseudo-marginalisation (with respect to both the forward-sampling and the backward-sampling operations), it cannot be described equivalently as a special instance of the original PaRIS (even in its general form given by Algorithm~\ref{alg:ideal:PaRIS}) operating on an extended space. Thus, there is no free lunch when it concerns the theoretical analysis of this scheme. Furthermore, in the case of fully dominated HMMs and when forward sampling is guided by the bootstrap filter, which was the setting in \cite{olsson:westerborn:2014b}, the conditional distribution of the particles given their ancestors (\ie, the marginal of $\rho_n$ in \eqref{eq:cond:instrumental:mixture} with respect to $x$) coincides, at any time point $n$, with the denominator of the backward kernel. Mathematically, this enables a cancellation that facilitates the analysis significantly. However, this simplification is not possible once the particle dynamics is guided by general proposal kernels, which is necessarily the case in Algorithm~\ref{alg:pm:PaRIS}. This complicates the analysis; see Remark~\ref{rem:non-trivial:extension} in the supplement for further details.

@@

We now focus on the asymptotic properties of Algorithm~\ref{alg:pm:PaRIS} and furnish, in Theorem~\ref{cor:clt:pseudo:marginal:paris} below, this scheme with a CLT. The stochastic stability of the algorithm is expressed by establishing, in Theorem~\ref{thm:variance:bound}, an $\mathcal{O}(n)$ bound on the asymptotic variances. First, let for $(x_n, x_{n + 1}) \in \Xset_n \times \Xset_{n + 1}$, 
\begin{multline} \label{eq:def:ukestvar}
\ukestvar{n}(x_n, x_{n + 1}) \eqdef 
\frac{1}{\wgtfuncmod{n}(x_n, x_{n + 1})} \int \{ \wgtfunc{n}(x_n, x_{n + 1}, z) - \wgtfuncmod{n}(x_n, x_{n + 1}) \}^2 \, \ukdist{n}(x_n, x_{n + 1}, \rmd z)
\end{multline}
denote the conditional relative variance of the (random) weight $\ewght{n + 1}{i}$ given $\epart{n}{\ind{n}{i}} = x_n$ and $\epart{n + 1}{i}= x_{n + 1}$. In addition, for each $n \in \nset$ and $m \in \intvect{0}{n}$, define the kernel  
\begin{equation} \label{eq:def:retrokmodmod}
    \retrokmodmod_{m, n}(x_m', \rmd x_{0:n}) \eqdef \delta_{x_m'}(\rmd x_m) \,  
    \tstatmod{m}(x_m, \rmd x_{0:m - 1})
    \prod_{\ell = m}^{n - 1} \ukmod{\ell}(x_\ell, \rmd x_{\ell + 1})
\end{equation}
on $\Xset_m \times \Xfd^n$ as well as the centered version 
\begin{equation} \label{eq:def:retrokmodmodnorm}
\retrokmodmodnorm_{m, n} h (x_m) \eqdef  \retrokmodmod_{m, n}(h - \postmod{0:n} h)(x_m) 
\end{equation}
on the same space. The following is the main result of this section. 
\begin{theorem} \label{cor:clt:pseudo:marginal:paris}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $\precpar \in \precparsp$, $\K \in \nsetpos$, and $h_n \in \bmaf{\Xfd^n}$, as $\N \to \infty$, 
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \postmod{0:n} h_n  \right) 
  \dlim \sigma_n(h_n) Z, 
$$
where $Z$ is standard normally distributed and 
\begin{equation} \label{eq:as:var:decomp}
\sigma^2_n (h_n) \eqdef \frac{\chi(\initwgtfunc \retrokmodmodnorm_{0, n} h_n)^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} + \sigma^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) + \sigma^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) 
\end{equation}
with
\begin{multline} \label{eq:non-recursive:as:var}
\sigma^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h_n) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m}(\varphi_m [\retrokmodmodnorm_{m + 1, n} h_n ]^2)}{(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}\\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\postmod{m} \adjfuncforward{m} \postmod{\ell} \ukmod{\ell} \{\bkmod{\ell}(\tstatmod{\ell} h_{\ell} + \addf{\ell} - \tstatmod{\ell + 1} h_{\ell +1})^2 \ukmod{\ell + 1, m}( \bkmod{m} \varphi_m [\ukmod{m + 1, n - 1} \1_{\Xset^n}]^2
)\}}{\K^{m - \ell + 1} (\postmod{\ell} \ukmod{\ell, m - 1} \1_{\Xset^m})(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}
\end{multline}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_\ell : \Xset_\ell \times \Xset_{\ell + 1} \to \rsetnn$. 
\end{theorem}

@@

Our analysis provides, as a by-product, also a CLT for the random-weight particle filter obtained by iterating the forward-sampling operation in Algorithm~\ref{alg:pm:SMC}; indeed, this result, which generalises \cite[Theorem~3]{fearnhead2008particle}, follows immediately by letting $f_n \equiv 0$ in Theorem~\ref{prop:clt:pseudo:marginal:paris} in the supplement, and we state it below for completeness.

\begin{proposition} \label{prop:clt:pseudo:marginal:filter}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal} and let $(\epart{n}{i}, \ewght{n}{i})_{i = 1}^\N$, $n \in \nset$, be generated by Algorithm~\ref{alg:pm:SMC}. Then for all $n \in \nset$, $\precpar \in \precparsp$, and $h \in \bmf{\Xfd_n}$, as $\N \to \infty$,
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} h(\epart{n}{i}) - \postmod{n} h \right) 
  \dlim \tilde{\sigma}_n(h) Z, 
$$
where $Z$ is standard normally distributed and
\begin{equation} \label{eq:as:var:pm:filter}
\tilde{\sigma}^2_n(h) \eqdef \frac{\chi\{ \initwgtfunc \ukmod{0} \cdots \ukmod{n - 1}(h - \postmod{n} h) \}^2}{(\chi \ukmod{0} \cdots \ukmod{n - 1} \1_{\Xset_n})^2} + \tilde{\sigma}^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h) + \tilde{\sigma}^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h) 
\end{equation}
with
\begin{equation} \label{eq:non-recursive:filter:as:var}
\tilde{\sigma}^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m}(\varphi_m [\ukmod{m + 1} \cdots \ukmod{n - 1} (h - \postmod{n} h)
]^2)}{(\postmod{m} \ukmod{m} \cdots \ukmod{n - 1} \1_{\Xset_n})^2} 
\end{equation}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_\ell : \Xset_\ell \times \Xset_{\ell + 1} \to \rsetnn$. 
\end{proposition}

@@

Importantly, Theorem~\ref{cor:clt:pseudo:marginal:paris} provides, as another by-product, also a CLT for the ideal PaRIS in Algorithm~\ref{alg:ideal:PaRIS}. Indeed, in the case where every $\ud{n}$ is tractable, we may set 
\begin{equation} \label{eq:ideal:case}
\ukest{n}{z}(x_n, x_{n + 1}) = \ud{n}(x_n, x_{n + 1})
\end{equation} 
for all $z$ and define $\ukdist{n}$ arbitrarily. This implies that the relative variance \eqref{eq:def:ukestvar} is identically zero, which eliminates the last term of \eqref{eq:as:var:decomp}. Thus, in this case the asymptotic variance is given by the first two terms of \eqref{eq:as:var:decomp}, but now induced by the original dynamics $(\uk{n})_{n \in \nset}$ (as \eqref{eq:ideal:case} implies that also $\udmod{n} = \ud{n}$). This result is formulated in the following corollary, where we have defined, for each $n \in \nset$ and $m \in \intvect{0}{n}$, the kernel  
\begin{equation} \label{eq:def:retrok}
    \retrok_{m, n}(x_m', \rmd x_{0:n}) \eqdef \delta_{x_m'}(\rmd x_m) \,  
    \tstat{n}(x_m, \rmd x_{0:m - 1})
    \prod_{\ell = m}^{n - 1} \uk{\ell}(x_\ell, \rmd x_{\ell + 1})
\end{equation}
on $\Xset_m \times \Xfd^n$ as well as the centered version 
$$
\retroknorm_{m, n} h (x_m) \eqdef  \retrok_{m, n}(h - \post{0:n} h)(x_m) 
$$
on the same space.

\begin{corollary} \label{cor:clt:ideal:paris}
For each $n \in \nset$, assume that the functions $\adjfuncforward{n}$ and  
$$
\wgtfuncideal{n} : \Xset_n \times \Xset_{n + 1} \ni (x_n, x_{n + 1}) \mapsto \frac{\ud{n}(x_n, x_{n + 1})}{\adjfuncforward{n}(x) \kissforward{n}{n}(x_n, x_{n + 1})} 
$$
are bounded and let $(\epart{n}{i}, \tstat[i]{n}, \ewght{n}{i})_{i = 1}^\N$, $n \in \nset$, be generated by Algorithm~\ref{alg:ideal:PaRIS}. Then for all $n \in \nset$, $\precpar \in \precparsp$, $\K \in \nsetpos$, and $h_n \in \bmaf{\Xfd^n}$, as $\N \to \infty$,
$$
\sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \post{0:n} h_n  \right) \dlim \bar{\sigma}_n (h_n) Z, 
$$
where $Z$ is standard normally distributed and
\begin{multline} \label{eq:as:var:ideal:PaRIS}
\bar{\sigma}^2_n (h_n) 
\eqdef \frac{\chi(\initwgtfunc \retroknorm_{0, n} h_n)^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} + \sum_{m = 0}^{n - 1} \frac{\post{m} \adjfuncforward{m} \post{m} \uk{m}(\wgtfuncideal{m} [\retroknorm_{m + 1, n} h_n ]^2)}{(\post{m} \uk{m, n - 1} \1_{\Xset^n})^2} \\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\post{m} \adjfuncforward{m} \post{\ell} \uk{\ell} \{\bkw{\ell}(\tstat{\ell} h_{\ell} + \addf{\ell} - \tstat{\ell + 1} h_{\ell +1})^2 \uk{\ell + 1, m}( \bkw{m} \wgtfuncideal{m} [\uk{m + 1, n - 1} \1_{\Xset^n}]^2
)\}}{\K^{m - \ell + 1} (\post{\ell} \uk{\ell, m - 1} \1_{\Xset^m})(\post{m} \uk{m, n - 1} \1_{\Xset^n})^2}. 
\end{multline}
\end{corollary}

@@

@@

Under \hypref{assum:strong:mixing} and \hypref{assum:bias:bound} we may establish the next theorem, which is the main result of this section. The proof is postponed to Section~\ref{sec:proofs} in the supplement.

@@

\begin{remark}
Consider now a parameterised version of the model, where the transition densities $(\ud{n ; \theta})_{k \in \nset}$ are indexed by some parameter $\theta$ belonging to some parameter space $\Theta$ being a subset of $\rset^d$. Assume further that all $(\ud{n ; \theta})_{n \in \nset}$ are differentiable with respect to $\theta$ and such that for all $n \in \nset$, $x_n \in \Xset$, and $h \in \bmf{\Xfd_{n + 1}}$,  
$$
\nabla_\theta \int \ud{n ; \theta}(x_n, x_{n + 1}) h(x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1}) = \int \nabla_\theta \ud{n ; \theta}(x_n, x_{n + 1}) h(x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1})
$$
and $\sup_{\theta \in \Theta} \int |\nabla_\theta \ud{n ; \theta}(x_n, x_{n + 1})| \, \mu_{n + 1}(\rmd x_{n + 1}) \leq c < \infty$ for some positive constant $c$. Then \hypref{assum:bias:bound} holds in the sense that for all $n \in \nset$, $h \in \bmf{\Xfd_{n + 1}}$, and $x_n \in \Xset$, 
$$
| \uk{n; \theta}h(x_n) - \uk{n; \theta'}h(x_n) | \leq c \| \theta - \theta' \| \| h \|_\infty
$$
(\emph{i.e.}, $\precpar = \| \theta - \theta' \|$ in this case). Finally, assume also that family satisfies \hypref{assum:strong:mixing} uniformly over the parameter space in the sense that for all $n \in \nset$, $(x_n, x_{n + 1}) \in \Xset^2$, and $\theta \in \Theta$, 
$
\udlow \leq \ud{n; \theta}(x_n, x_{n + 1}) \leq \udup. 
$  
Then Theorem~\ref{thm:bias:bound} provides a positive constant $d$ such that for all $n \in \nset$,  
$$
\big| \post{0:n ; \theta} h_n -  \post{0:n ; \theta'} h_n \big|  
\leq d n \| \theta - \theta' \| \sup_{k \in \nset} \| \addf{k} \|_\infty. 
$$
This extends previous results on the uniform continuity of the filter distribution (see, \eg, \cite{papavasiliou:2006,legland:oudjane:2004}) to smoothing of additive state functionals.  
\end{remark}
